Quantitative Auditing of AI Fairness with Differentially Private Synthetic Data

### 2025-01-19 paper/paper.tex
Cleared TODOs

Fairness in machine learning has become an increasingly important topic as AI systems are widely used across various sectors, including healthcare and criminal justice. These systems can significantly influence outcomes that impact individuals and communities.

@@

The key idea is that no matter how much background knowledge an attacker has about the data, they cannot determine with high confidence whether a specific individual is included in the dataset. This is quantified by some parameters which control the amount of noise introduced. Smaller parameters provide stronger privacy but can reduce the accuracy of the analysis, while larger parameters allow for more precise results but with weaker privacy guarantees. The challenge is to carefully calibrate the parameters to ensure that privacy is maintained without overly compromising the utility of the data.

@@

% TODO: add to evaluation
% TODO: add to conclusion

@@

The average of the differences between the fairness measure values of the original and synthetic datasets is below 0.1 for all datasets. This indicates that synthetic data is a good approximation of the original data in terms of fairness.

@@

In the Adult experiment, Conditional Use Accuracy Equality (True Positive) has an error of around 0.16. In the COMPAS experiment, both Conditional Use Accuracy Equality measures had an error over 0.1. In the Diabetes experiment, Conditional Use Accuracy Equality (True Negative) has an error much higher than other measures.

@@

Experiments conducted on multiple real-world datasets and models show that synthetic data derived from real data retains essential fairness properties. This ensures that any biases present in the original data are captured and reflected in the synthetic counterpart.

>>>>

Fairness in machine learning has become an increasingly important topic as AI systems are widely used across various sectors, including healthcare and criminal justice. These AI systems can significantly influence outcomes that impact individuals and communities.

@@

The key idea is that no matter how much background knowledge an attacker has about the data, they cannot determine with high confidence whether a specific individual is included in the dataset. This is quantified by some parameters which control the amount of noise introduced; smaller parameters provide stronger privacy but can reduce the accuracy of the analysis, while larger parameters allow for more precise results but with weaker privacy guarantees. The challenge is to carefully calibrate the parameters to ensure that privacy is maintained without overly compromising the utility of the data.

@@

@@

The average of the differences between the fairness measure values of the original and synthetic datasets is below $0.1$ for all datasets across all models. Judging from different metrics, datasets, and models, this indicates that synthetic data is a good approximation of the original data in terms of fairness.

@@

In the Adult random forest experiment, Conditional Use Accuracy Equality (True Positive) has an error of around 0.16. In the COMPAS random forest experiment, both Conditional Use Accuracy Equality measures had an error over 0.1. In the Diabetes random forest experiment, Conditional Use Accuracy Equality (True Negative) has an error much higher than other measures.

@@

Experiments conducted on multiple real-world datasets and across different models show that synthetic data derived from real data retains essential fairness properties. This ensures that any biases present in the original data are captured and reflected in the synthetic counterpart.
